---
title: "Open Prescribing Data Checks: Atorvastatin"
format: 
  html:
    fontsize: 11px
    mainfont: "Arial"
---

### Background

```{r}
#| label: load-packages
#| include: false

library(dplyr)
library(readr)
library(here)
library(glue)
library(tidyr)
library(stringr)
library(scales) ## format label_number
library(gt) ## gt table
library(ggplot2)
library(fpp2) # Forecasting: Principles and Practice" (2nd Edition)
library(lubridate) # interval
```

```{r}
#| label: import-data
#| include: false

source(here("analysis","design.R"))

label_number_n <- label_number(1, 1, big.mark=",")
label_perc_n <- label_percent(accuracy =.0001)


df_hscic_stp <- read_csv(here::here(directory_name,"prac_stps.csv")) 

###  
df_hscic_ccgs <- read_csv(here::here(directory_name,"prac_ccgs.csv")) %>%
  select("code","name","stp_id", "regional_team_id","org_type") 

df_hscic_practice_statistics <- read_csv(here::here(directory_name,"practice_statistics.csv"))

df_hscic_practices <- read_csv(here::here(directory_name,"practices.csv"))

df_hscic_normalised_prescribing <- read_csv(here::here(directory_name,"normalised_prescribing.csv.gz"))

### list size (restricted to type 4 practices)
df_stats_prac <- df_hscic_practice_statistics %>%
  left_join(df_hscic_practices,by=join_by(practice==code)) %>%
  filter(setting==4) %>%
  select(month,total_list_size,pct_id,practice,name,ccg_id,pcn_id,status_code)


## list size (not restricted to type 4 practices)
df_stats_prac_full <- df_hscic_practice_statistics %>%
  left_join(df_hscic_practices,by=join_by(practice==code)) %>%
  select(month,total_list_size,pct_id,practice,name,ccg_id,pcn_id,status_code)

## atorvastatin at regional level 
df_atorvastatin <- read_csv(here::here(directory_name,"atorvastatin_list.csv"))

## atorvastatin at practice level (includes stp,pct, bnf_name from hscic.normalised_prescribing)
df_atorvastatin_prac <- read_csv(here::here(directory_name,"atorvastatin_practice.csv.gz"))

### regions from region team
df_regions <- read_csv(here::here("data","NHS_England_Names_and_Codes_in_England.csv")) %>%
  rename(regional_team = NHSER24CDH ,region = NHSER24NM) %>%
  select(regional_team,region)

regions<-unique(df_regions$regional_team)

df_atorvastatin_prac_region <- df_atorvastatin_prac %>% 
  left_join(df_stats_prac) %>%
  mutate(month =as.Date(month)) %>% 
  ungroup()

start_date_ator <- min(df_atorvastatin_prac$month)

end_date_ator <- max(df_atorvastatin_prac$month)

study_months <- interval(start_date_ator, end_date_ator ) %/% months(1) +1
```

NHS Digital publishes monthly and annual prescribing datasets from the NHS Business Services Authority, along with static reports on prescribing trends. However, this does not allow readers to interrogate topics of interest in detail, and the large datasets can be complex to manage.

OpenPrescribing.net, an online and publicly accessible tool, to help users explore highly granular NHS primary care prescribing open data. It is widely used with over 15,000 unique users each month. Its users are predominantly from within the NHS, but industry and patient groups are also well represented.

In order to aid the interpretation of work using the OpenPrescribing dataset to investigate atorvastatin prescribing behaviors I looked at some of the key data quality issues.

### Methods

#### Study design

The data used here is downloaded from [**OpenPrescribing**](https://openprescribing.net/) using BigQuery via the notebooks in the [notebook directory](https://github.com/andrewscolm/op_indicator_saturation/blob/master/notebooks).

5 tables are queried.

-   `ebmdatalab.hscic.normalised_prescribing`

```{r}
#| label: table-normalised-prescribing
#| echo: false
#| warning: false

normalised_prescribing <- tibble(column=c("regional_team","stp","pct","practice","bnf_code","bnf_name","items","net_cost","actual_cost","quantity","month"),description=c("Regional team identifier","Sustainability and Transformation Partnership code","Primary Care Trust code","Practice code","	British National Formulary (BNF) code","British National Formulary (BNF) drug name","Number of items (one item = one appearance on a prescription)","Cost of the presentation according to the Drug Tariff","Actual cost including adjustments for bulk purchases, out of pocket expenses, etc. (the ‘cost’ field typically used in queries)","Number of pills/grams/millilitres/dressings/ampoules prescribed","Prescribing month"))

gt(normalised_prescribing) %>%
  tab_options(table.font.size = 10,
  data_row.padding = px(1)) %>% 
    tab_style(
      style = cell_borders(
      sides = "all",
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = list(cells_body(), cells_column_labels())) 

```

-   `ebmdatalab.hscic.practices`

```{r}
#| label: table-hscic.practices
#| echo: false
#| warning: false

hscic.practices <- tibble(column=c("code","name","address1","address2","address3","address4","address5","postcode","location","ccg_id","pcn_id","setting","close_date","join_provider_date","leave_provider_date","open_date","status_code"),description=c("Practice code","Practice name","Address line 1","Address line 2","Address line 3","Address line 4 (town/city)","Address line 5","Practice postcode","Location of practice in WGS84 format","Clinical Commissioning Group identifier","Primary Care Network identifier","Practice setting code (see hscic.prescribing_setting)","Date practice closed","Date practice joined CCG","Date practice left CCG (or closed)","Date practice opened","Practice status code (A=Active, B=Retired, C=Closed, D=Dormant, P=Proposed, U=Unknown)"))

gt(hscic.practices ) %>%
  tab_options(table.font.size = 10,
  data_row.padding = px(1)) %>% 
    tab_style(
      style = cell_borders(
      sides = "all",
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = list(cells_body(), cells_column_labels())) 

```

-   `ebmdatalab.practice_statistics`

```{r}
#| label: table-practice_statistics
#| echo: false
#| warning: false

practice_statistics <- tibble(column=c("month","male_*","female_*","total_list_size","astro_pu_cost","astro_pu_items","star_pu","pct_id","practice"),description=c("The month for which practice statistics are recorded","Count of male patients aged *","Count of female patients aged *","Total count of patients registered at the practice","ASTRO-PU (Age, Sex, and Temporary Resident Originated Prescribing Units) cost weighting","ASTRO-PU items weighting","STAR-PU (Specific Therapeutic Group Age-sex weighting Related Prescribing Unit)","Primary Care Trust identifier","Practice code"))

gt(practice_statistics ) %>%
  tab_options(table.font.size = 10,
  data_row.padding = px(1)) %>% 
    tab_style(
      style = cell_borders(
      sides = "all",
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = list(cells_body(), cells_column_labels())) 

```

-   `ebmdatalab.hscic.ccgs`

```{r}
#| label: table-hscic.ccgs
#| echo: false
#| warning: false

hscic.ccgs <- tibble(column=c("code","name","ons_code","org_type","open_date","close_date","address","postcode","stp_id","regional_team_id")
          ,description=c("The organisation (CCG / PCT / Unknown) code","The organisation name","The ONS code (pct only)","The organisation type (CCG / PCT / Unknown)","The date the organisation opened (CCG or Unknown only)","The date the organisation closed (CCG or Unknown only)","The organisation address (CCG or Unknown only)","The organisation postcode (CCG or Unknown only)","Sustainability and Transformation Partnership code (CCG only)","Regional team identifier (CCG or Unknown only)"))

gt(hscic.ccgs ) %>%
  tab_options(table.font.size = 10,
  data_row.padding = px(1)) %>% 
    tab_style(
      style = cell_borders(
      sides = "all",
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = list(cells_body(), cells_column_labels())) 

```

-   `hscic.bnf`

```{r}
#| label: table-hscic.bnf
#| echo: false
#| warning: false

hscic.bnf <- tibble(column=c("year_month","chapter","chapter_code","section","section_code","para","para_code","subpara","subpara_code","chemical","chemical_code","product","product_code","presentation","presentation_code"),description=c("year_month","The chemical’s BNF chapter","The chemical’s BNF chapter code","The chemical’s BNF section","The chemical’s BNF section code","The chemical’s BNF para","The chemical’s BNF para code","The chemical’s BNF subpara","The chemical’s BNF subpara code","The chemical’s BNF chemical","The chemical’s BNF chemical code","The chemical’s BNF product","The chemical’s BNF product code","The chemical’s BNF presentation","The chemical’s BNF presentation code"))


gt(hscic.bnf ) %>%
  tab_options(table.font.size = 10,
  data_row.padding = px(1)) %>% 
    tab_style(
      style = cell_borders(
      sides = "all",
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = list(cells_body(), cells_column_labels())) 

```

-   `ebmdatalab.hscic.prescribing_setting`

```{r}
#| label: table-prescribing_setting
#| echo: false
#| warning: false

prescribing_setting <- tibble(setting=c("0","1","2","3","4","8","9","10","11","12","13","14","15","16","17","18","19","20","21","22","24","25","26","27"),
        description=c("Other","WIC Practice","OOH Practice","WIC + OOH Practice","GP Practice","Public Health Service","Community Health Service","Hospital Service","Optometry Service","Urgent & Emergency Care","Hospice","Care Home / Nursing Home","Border Force","Young Offender Institution","Secure Training Centre","Secure Childrens Home","Immigration Removal Centre","Court","Police Custody","Sexual Assault Referral Centre (SARC)","Other Justice Estate","Prison","PCN","Independent Pharmacy"))

gt(prescribing_setting ) %>%
  tab_options(table.font.size = 10,
  data_row.padding = px(1)) %>% 
    tab_style(
      style = cell_borders(
      sides = "all",
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = list(cells_body(), cells_column_labels())) 

```

#### Study population

Practices which prescribed atorvastatin during the study period of `{r} format(start_date_ator , "%B %Y")` - `{r} format(end_date_ator , "%B %Y")`

#### Measures

Items: Items are a count of the number of times a medicine has been prescribed. A prescription item refers to a single supply of a medicine, dressing or appliance prescribed on a prescription form. Item figures do not provide any indication of the length of treatment or quantity of medicine prescribed.

List size: The list size is the number of patients registered to a practice on a given date.

#### Statistical methods

I investigated:

Practices contributing to every month: Data is extracted monthly for each practice. Practices may open, close, or stop prescribing atorvastatin during the study period. In order to assess the impact of restricting analysis to practices which contribute to every month of the study period I investigated the proportion of practices contributing to every month.

Missing practice_statistics information The list size, Sub ICB Location, and practice name come from the practice_statistics table. This table does not include information for all of the practices for every month.

Practices with no associated list size I looked at i) the proportion of items that were prescribed from a practice with no list size. ii) whether practices were missing list size throughout the study, after a period of having a list size, or before a period of having a list size.

Identifying atypical practices: The ebmdatalab.hscic.practices table is used to restrict the datasets to GP settings. However, this does not always perfectly restrict to typical GP practices. Further work needs to be applied in order to identify atypical practices such as special allocation scheme, specialist, urgent, asylum, pilot or nursing home practices. I attempted to identify the practice type based on the practice name.

Practices with no 20mg atorvastatin prescribing: Practices are only included if they have a prescription for any strength of atorvastatin for the month in question. I investigated how many practices were dropped when restricting to practices with 20mg atorvastatin prescription.

Practices which changed region during the study period: Sub-ICB locations (SICBL) can be identified via 3 different columns in three different datasets. The ccg_id in hscic.practices is static and shows the SICBL as it is now. The pct_id column in hscic.practice-statistics is also static. The pct column in hscic.normalised-prescribing shows the SICBL as it was at the time. I investigated practices that changed region during the study period.

List size changes for practices that either initiate or stop prescribing atorvastatin during the study period.

### Results

#### Practices contributing to every month.

```{r}
#| label: all-month-filter
#| include: false

# count number of times each practice contributes 
prac_count<-df_atorvastatin_prac %>% 
    group_by(practice,month) %>% 
    summarise(practice =last(practice)) %>%
    count(practice) ## 6858

atorvastatin_practices <- df_atorvastatin_prac %>%
  group_by(practice) %>%
  summarise(practice=last(practice)) 
  
    
n_atorvastatin_practices <- atorvastatin_practices %>%
  nrow()  ## 6858


if( nrow(prac_count) != n_atorvastatin_practices){ print("Practice count does not match with above method")}

# do we lose practices in the filter?
prac_count_all_months<-prac_count %>% 
    filter(n==study_months) # 6,164

 id_dropped_count_all_months <- prac_count %>%
  anti_join(prac_count_all_months %>% select(practice)) %>%
  select(practice)

dropped_count_all_months <- df_atorvastatin_prac_region %>% 
  inner_join(id_dropped_count_all_months) 

dropped_count_all_months_items <-  sum(dropped_count_all_months$items)
  
df_dropped_count_all_months_list_size <- dropped_count_all_months %>%
  group_by(practice,month) %>%
  summarise(total_list_size = mean(total_list_size ,na.rm = T)) %>%
  group_by(practice) %>%
  summarise(total_list_size = sum(total_list_size ,na.rm = T))

dropped_count_all_months_list_size <-sum(df_dropped_count_all_months_list_size$total_list_size)

```

`{r} label_number_n(nrow(prac_count_all_months))` of `{r} label_number_n(nrow(prac_count))` practices prescribed atorvastatin throughout the study period. `{r} label_number_n(nrow(prac_count)-nrow(prac_count_all_months))` practices will be lost by restricting to practices which prescribe every month of the study. These practices contribute `{r} label_number_n(dropped_count_all_months_items)` prescribed items and a total list size of `{r} label_number_n(dropped_count_all_months_list_size)` throughout the study period.

#### Missing practice_statistics information

```{r}
#| label: miss-sicbl
#| echo: false
#| warning: false

### how many practices are missing sicbl code  
miss_ccg <- df_atorvastatin_prac  %>% 
  left_join(df_stats_prac) %>% 
  filter(is.na(ccg_id)) # 4,415 lines with missing sicbl

miss_ccg_id <- miss_ccg %>%
  group_by(practice) %>%
  summarise(practice=last(practice)) #399 practices missing ccg (means missing list size)



```

There are `{r} label_number_n(nrow(miss_ccg_id))` practices missing at least one month from the `practice_statistics` table, contributing `{r} label_number_n(nrow(miss_ccg))` months of data. Practices are missing across all regions.

```{r}
#| label: table-miss-ccg-by-region
#| echo: false
#| warning: false
miss_ccg_by_region <- miss_ccg %>%
  group_by(practice) %>%
  summarise(practice=last(practice),
            regional_team=last(regional_team)) %>% #399 practices missing ccg (means missing list size)
  left_join(df_regions) %>%
  group_by(region) %>%
  count() %>%
  ungroup()

if( sum(miss_ccg_by_region$n) != nrow(miss_ccg_id)){ print("Practice count does not match with above method")}

gt(miss_ccg_by_region) %>%
  tab_options(table.font.size = 10,
  data_row.padding = px(1)) %>% 
    tab_style(
      style = cell_borders(
      sides = "all",
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = list(cells_body(), cells_column_labels())) 

```

#### Missing list sizes

```{r}
#| label: miss-list-size
#| echo: false
#| warning: false

#### practices with no list size
df_atorvastatin_nolist <- df_atorvastatin_prac_region %>%
  filter(is.na(total_list_size)) %>%
  mutate(total = sum(items))  # 25,482 items prescribed where there is no list size

n_df_atorvastatin_nolist <- df_atorvastatin_nolist %>% 
    select(total) %>% 
    unique() %>%
    unlist()

df_atorvastatin_list<-df_atorvastatin_prac_region %>%
  filter(!is.na(total_list_size)) %>%
  mutate(total = sum(items)) ## 338,154,101 items prescribed where there is a list size

n_df_atorvastatin_list <- df_atorvastatin_list %>% 
    select(total) %>% 
    unique() %>%
    unlist()

region_missing<-df_atorvastatin_nolist %>% 
  select(regional_team) %>%
  unique() #  all regions have missing list sizes


df_atorvastatin_nolist_id <- df_atorvastatin_nolist %>% 
  select(practice) %>%
  unique() # 399 practices with missing list size

# df_atorvastatin_prac_region %>%
 # right_join(df_atorvastatin_nolist_id)  ### some missing at start, some at end, some all the way
```

The `{r} nrow(df_atorvastatin_nolist_id)` practices missing from the `practice_statistics` table all contribute prescribing data with no list size (Appendix 2). `{r} label_number_n(n_df_atorvastatin_nolist)` out of `{r} label_number_n(n_df_atorvastatin_list + n_df_atorvastatin_nolist)` (`{r} label_perc_n(n_df_atorvastatin_nolist/(n_df_atorvastatin_list + n_df_atorvastatin_nolist))`) items were prescribed from a practice with no associated list size. I identified practices that prescribed items and had:

-   no list size throughout the study

-   a period with no list size preceding a period with a list size

-   a period with a list size preceding a period with no list size

-   a period with a list size bookended by periods without a list size

-   a period without a list size bookended by periods with a list size

```{r}
#| label: figure-missing-list-sizes
#| echo: false
#| warning: false
#| fig.width: 25
#| fig.height: 12.5

### Nice graph of practices dropping in and out of list size

missing_list_size_prac <- df_atorvastatin_prac_region %>%
  group_by(month,practice) %>%
  summarise(
    items=sum(items),
    total_list_size=mean(total_list_size, na.rm = TRUE),
    regional_team= last(regional_team)) %>%
  right_join(df_atorvastatin_nolist_id) %>% 
  group_by(month,practice,) %>%
  summarise(list_size= factor(is.na(total_list_size)*1,levels =c(1,0),labels =c("items\nwith\nno list size","items\n&\nlist size")),
            regional_team= last(regional_team)) %>%
              ungroup()


for (regional in regions){
  assign(glue("plot_missing_list_size_prac_{regional}"), missing_list_size_prac %>%
  filter(regional_team==regional) %>%
  ggplot(aes(x=month, y=practice, fill=list_size)) +
  geom_tile() +
    ggtitle(df_regions %>% filter(regional_team==regional)%>%select(region)))
}

plot_missing_list_size_prac_NA<-missing_list_size_prac %>%
  filter(is.na(regional_team)) %>%
  ggplot(aes(x=month, y=practice, fill=list_size)) +
  geom_tile() +
  ggtitle("NA")

A<-ggplot() +theme_minimal()

plot_missing_list_size_prac <- patchwork::wrap_plots(
                      plot_missing_list_size_prac_Y56,
                      plot_missing_list_size_prac_Y58,
                      plot_missing_list_size_prac_Y59,
                      plot_missing_list_size_prac_Y60,
                      plot_missing_list_size_prac_Y61,
                      plot_missing_list_size_prac_Y62,
                      plot_missing_list_size_prac_Y63,
                      ncol = 7,
                      guides = "collect")
#                      heights = c(1,0.1))

ggsave(
  filename = here::here(
    "output",
    "data_check",
    "atorvastatin",
    "missing_list_size",
    glue("list_size_missingness.png")),
  plot_missing_list_size_prac,
  dpi = 800,
  width = 60,
  height = 30,
  units = "cm"
)

plot_missing_list_size_prac
```


```{r}
#| label: missing-list-sizes-patterns
#| echo: false
#| warning: false
#| fig.width: 25
#| fig.height: 12.5

missing_list_size_prac_first <- missing_list_size_prac %>%
  arrange(month) %>%
  group_by(practice) %>%
  slice_min(month) %>%
  select(practice,list_size)

missing_list_size_prac_last <- missing_list_size_prac %>%
  arrange(month) %>%
  group_by(practice) %>%
  slice_max(month) %>%
  select(practice,list_size)


missing_list_size_prac_change <- missing_list_size_prac %>%
  group_by(practice) %>%
  select(practice,list_size) %>%
  unique() %>% 
  count(practice)

table_only_missing_list_size <- missing_list_size_prac_first %>%
  filter(list_size=="items\nwith\nno list size") %>%
  inner_join(missing_list_size_prac_change %>%
  filter(n==1))

table_gains_list_size <- missing_list_size_prac_first %>%
  filter(list_size=="items\nwith\nno list size") %>%
  inner_join(missing_list_size_prac_last %>%
  filter(list_size=="items\n&\nlist size"),join_by(practice))

table_loses_list_size <- missing_list_size_prac_first %>%
  filter(list_size=="items\n&\nlist size") %>%
  inner_join(missing_list_size_prac_last %>%
    filter(list_size=="items\nwith\nno list size"),join_by(practice))

table_bookended_missing_list_size <- missing_list_size_prac_first %>%
  filter(list_size=="items\nwith\nno list size") %>%
  inner_join(missing_list_size_prac_last %>%
    filter(list_size=="items\nwith\nno list size"),join_by(practice)) %>%
  inner_join(missing_list_size_prac_change %>%
  filter(n==2))

table_bookended_with_list_size <- missing_list_size_prac_first %>%
  filter(list_size=="items\n&\nlist size") %>%
  inner_join(missing_list_size_prac_last %>%
    filter(list_size=="items\n&\nlist size"),join_by(practice)) %>%
  inner_join(missing_list_size_prac_change %>%
  filter(n==2))

list_sizeless<-table_only_missing_list_size %>% select(practice) %>% mutate(pattern="no list size throughout") %>%
  bind_rows(table_gains_list_size  %>% select(practice) %>% mutate(pattern="initially no list size" )) %>%
  bind_rows(table_loses_list_size  %>% select(practice) %>% mutate(pattern="initial list size" )) %>%
  bind_rows(table_bookended_missing_list_size  %>% select(practice) %>% mutate(pattern="bookended by missing list size" )) %>%
  bind_rows(table_bookended_with_list_size  %>% select(practice) %>% mutate(pattern="bookended by list size" )) %>%
  left_join(df_stats_prac %>% group_by(practice) %>% summarise(name=last(name))) %>%
  ungroup() %>%
  gt() %>%
  tab_options(table.font.size = 10,
  data_row.padding = px(1)) %>%
    tab_style(
      style = cell_borders(
      sides = "all",
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = list(cells_body(), cells_column_labels())) %>% tab_style(
      style = cell_borders(
      sides = "all",
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = list(cells_body(), cells_column_labels())) 

```


##### no list size throughout the study
`{r} nrow(table_only_missing_list_size)` practices have prescribing data with no list size throughout the study period. An example of this pattern can be seen in practice `Y02666` below:

```{r}
#| label: figure-missing-list-sizes-always-missing
#| echo: false
#| warning: false
#| fig.width: 5
#| fig.height: 1
plot_prac_2<-missing_list_size_prac %>%
  filter(practice=="Y02666") %>%
  ggplot(aes(x=month, y=practice, fill=list_size)) +
  geom_tile(fill = "#00BFC4") +
  xlim(as.Date("2020-01-01"),as.Date("2025-08-01"))

plot_prac_2

ggsave(
  filename = here::here(
    "output",
    "data_check",
    "atorvastatin",
    "missing_list_size",
    glue("plot_prac_2.png")),
  plot_prac_2,
  dpi = 800,
  width = 14,
  height = 8,
  units = "cm"
)
```

Practice `Y02666` is Salisbury walk-in centre. As a walk in centre there is no fixed patient list. However, it is able to prescribe atorvastatin.

##### a period with no list size preceding a period with a list size 
`{r} nrow(table_gains_list_size)` practices have only a list size for a period before they have both prescribing and list size. An example of this pattern can be seen in practice `Y07040` below:

```{r}
#| label: figure-missing-list-sizes-lead-in
#| echo: false
#| warning: false
#| fig.width: 5
#| fig.height: 1
#| 
plot_prac_1<-missing_list_size_prac %>%
  filter(practice=="Y07040") %>%
  ggplot(aes(x=month, y=practice, fill=list_size)) +
  geom_tile() +
  xlim(as.Date("2020-01-01"),as.Date("2025-08-01"))
plot_prac_1

ggsave(
  filename = here::here(
    "output",
    "data_check",
    "atorvastatin",
    "missing_list_size",
    glue("plot_prac_1.png")),
  plot_prac_1,
  dpi = 800,
  width = 14,
  height = 8,
  units = "cm"
)

```

Practice `Y07040` is Medicus Select Care. This is an example of a Special Allocation Scheme (SAS) practice. The SAS program provides primary care services to patients who have been removed from their regular GP practice list due to violent or aggressive behavior that was reported to the police. This is not a typical GP practice.

##### a period with a list size preceding a period with no list size

`{r} nrow(table_loses_list_size)` practices have prescribing data with no list size for a period after they have both prescribing and list size. An example of this pattern can be seen in practice `L83642` below:

```{r}
#| label: figure-missing-list-sizes-end-out
#| echo: false
#| warning: false
#| fig.width: 5
#| fig.height: 1
plot_prac_3<-missing_list_size_prac %>%
  filter(practice=="L83642") %>%
  ggplot(aes(x=month, y=practice, fill=list_size)) +
  geom_tile() +
  xlim(as.Date("2020-01-01"),as.Date("2025-08-01"))
plot_prac_3

ggsave(
  filename = here::here(
    "output",
    "data_check",
    "atorvastatin",
    "missing_list_size",
    glue("plot_prac_3.png")),
  plot_prac_3,
  dpi = 800,
  width = 14,
  height = 8,
  units = "cm"
)

```

Practice `L83642` is Estover Surgery. Estover surgery closed however it is possible that, due to electronic repeat dispensing (eRD), atorvastatin may still have been dispensed for 6-12 months after closing.


##### a period with a list size bookended by periods without a list size
There is only one practice (Y07274) bookended by periods with no list size. 

```{r}
#| label: figure-missing-list-sizes-bookend-without
#| echo: false
#| warning: false
#| fig.width: 5
#| fig.height: 1
plot_prac_5<-missing_list_size_prac %>%
  filter(practice=="Y07274") %>%
  ggplot(aes(x=month, y=practice, fill=list_size)) +
  geom_tile() +
  xlim(as.Date("2020-01-01"),as.Date("2025-08-01"))
plot_prac_5

ggsave(
  filename = here::here(
    "output",
    "data_check",
    "atorvastatin",
    "missing_list_size",
    glue("plot_prac_5.png")),
  plot_prac_5,
  dpi = 800,
  width = 14,
  height = 8,
  units = "cm"
)

```

Practice Y07274 is St Nicolas Medical Centre, a now dormant GP Practice.



##### a period without a list size bookended by periods with a list size

`{r} nrow(table_bookended_with_list_size)` practices have a period of prescribing with no list size bookended by periods with list size. An example of this pattern can be seen in practice `L81094` below:

```{r}
#| label: figure-missing-list-sizes-bookended
#| echo: false
#| warning: false
#| fig.width: 5
#| fig.height: 1
plot_prac_3<-missing_list_size_prac %>%
  filter(practice=="L81094") %>%
  ggplot(aes(x=month, y=practice, fill=list_size)) +
  geom_tile() +
  xlim(as.Date("2020-01-01"),as.Date("2025-08-01"))
plot_prac_3

ggsave(
  filename = here::here(
    "output",
    "data_check",
    "atorvastatin",
    "missing_list_size",
    glue("plot_prac_4.png")),
  plot_prac_3,
  dpi = 800,
  width = 14,
  height = 8,
  units = "cm"
)

```

Practice `L81094` is The Merrywood Practice which closed in February 2025. The list sizes before February 2024 ranged from 6673-7321. The list size for May 2024 (after the period with no list size was 1). This may indicate that the practice effectively closed before the official closure date. 




#### Identifying atypical practices

I used the following terms to try to identify each practice type from the practice name:

-   special allocation scheme practice
    -   "SAS"
    -   "Special Allocation"
    -   "Select care"
    -   "SPCL"
-   specialist practice
    -   "Specialist"
    -   "LARC"
-   urgent care
    -   "urgent"
-   asylum practice
    -   "asylum"
-   pilot practice
    -   "Pilot"
-   nursing home practice
    -   "nurs"
    -   "one care"
-   Walk-in centres
    -   "walk-"
    -   "walk in"
    -   "unregistered"
-   Outside standard hours
    -   "7 day access"
    -   "EA hub"


```{r}
#| label: settings
#| include: false

setting_sas <- df_stats_prac %>%
  mutate(name=str_to_lower(name)) %>%
 filter(str_detect(name, 'sas|special allocation|select care|spcl')) %>%
 group_by(practice, name) %>%
 summarise(name = last(name),
 type= "special allocation scheme")

setting_specialist <- df_stats_prac %>%
  mutate(name=str_to_lower(name)) %>%
 filter(str_detect(name, 'specialist|larc')) %>%
 group_by(practice, name) %>%
 summarise(name = last(name),
 type= "specialist")

setting_urgent <- df_stats_prac %>%
  mutate(name=str_to_lower(name)) %>%
 filter(str_detect(name, 'urgent')) %>%
 group_by(practice, name) %>%
 summarise(name = last(name),
 type= "urgent care")

setting_asylum <- df_stats_prac %>%
  mutate(name=str_to_lower(name)) %>%
 filter(str_detect(name, 'asylum')) %>%
 group_by(practice, name) %>%
 summarise(name = last(name),
 type= "asylum")

setting_pilot <- df_stats_prac %>%
  mutate(name=str_to_lower(name)) %>%
 filter(str_detect(name, 'pilot')) %>%
 group_by(practice, name) %>%
 summarise(name = last(name),
 type= "pilot")

setting_nursing <- df_stats_prac %>%
  mutate(name=str_to_lower(name)) %>%
 filter(str_detect(name, 'nurs|one care')) %>%
 group_by(practice, name) %>%
 summarise(name = last(name),
 type= "nursing home")

setting_walk <- df_stats_prac %>%
  mutate(name=str_to_lower(name)) %>%
 filter(str_detect(name, 'walk-|walk in|unregistered')) %>%
 group_by(practice, name) %>%
 summarise(name = last(name),
 type= "walk-in")

 setting_outside_hours <- df_stats_prac %>%
  mutate(name=str_to_lower(name)) %>%
 filter(str_detect(name, '7 day access|ea hub')) %>%
 group_by(practice, name) %>%
 summarise(name = last(name),
 type= "outside-hours")

### get highest total_list_size
high_list_size <- df_stats_prac %>%
 group_by(name) %>%
 summarise(name =str_to_lower(last(name)),
 highest_list_size = max(total_list_size ,na.rm = T))

df_setting <- setting_sas %>%
  bind_rows(setting_specialist) %>%
  bind_rows(setting_urgent) %>%
  bind_rows(setting_asylum) %>%
  bind_rows(setting_pilot) %>%
  bind_rows(setting_nursing) %>%
  bind_rows(setting_walk) %>%
  bind_rows(setting_outside_hours) %>%
  left_join(high_list_size) %>% 
  ungroup()



```

In total, I identified `{r} nrow(df_setting)` atypical practices. This comprised `{r} nrow(setting_sas)` special allocation scheme practices, `{r} nrow(setting_specialist)` specialist scheme practice,`{r} nrow(setting_urgent)` urgent care practice, `{r} nrow(setting_asylum)` asylum practice, `{r} nrow(setting_pilot)` pilot scheme practice and `{r} nrow(setting_nursing)` nursing home practice. Filtering practices by a minimum list size is sometimes used to remove these atypical practices. The following table contains the maximum list size of each identified atypical practice during the study period.

```{r}
#| label: table_settings
#| echo: false
#| warning: false



df_setting %>%
  gt() %>%
  cols_label(highest_list_size = "maximum list size") %>%
  tab_options(table.font.size = 10,
  data_row.padding = px(1)) %>%
    tab_style(
      style = cell_borders(
      sides = "all",
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = list(cells_body(), cells_column_labels())) %>% tab_style(
      style = cell_borders(
      sides = "all",
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = list(cells_body(), cells_column_labels())) 


```

#### Practices with no 20mg atorvastatin prescribing.

```{r}
#| label: 20mg-filter
#| include: false

# filter to 20
df_atorvastatin_prac_20 <- df_atorvastatin_prac %>%
  filter(str_detect(bnf_name, '20')) 

# do we lose practices in the filter?
atorvastatin_20_practices <- df_atorvastatin_prac_20 %>%  
  group_by(practice) %>%
  summarise(practice=last(practice)) 


n_atorvastatin_20_practices <- atorvastatin_20_practices %>%
  nrow() ## 6845, (n_atorvastatin_practices 6858)

### identify lost practices
id_ator_lost_list <- atorvastatin_practices %>% 
  anti_join(atorvastatin_20_practices)

```

Over the entire study period there are `{r} label_number_n(n_atorvastatin_practices)` practices prescribing atorvastatin, with `{r} label_number_n(n_atorvastatin_20_practices)` practices prescribing 20mg strength. We lose `{r} n_atorvastatin_practices - n_atorvastatin_20_practices` practices by restricting to 20mg prescriptions. The practices dropped are as follows:

```{r}
#| label: table-20mg-filter
#| echo: false
#| warning: false

## table of practices lost when filtered to 20mg
df_ator_lost_list <- id_ator_lost_list %>%
  left_join(df_stats_prac_full) %>%
  group_by(practice) %>%
  summarise(prac_name = last(name),
            ccg_id=last(ccg_id)) %>%
  left_join(df_hscic_ccgs,by=join_by(ccg_id ==code)) %>%
  left_join(df_regions,by=join_by(regional_team_id ==regional_team )) %>%
  select(practice,prac_name,ccg_id,name,regional_team_id,region) %>%
  ungroup()

gt(df_ator_lost_list) %>%
  sub_missing(missing_text = "-") %>%
  cols_label(
    practice = "Practice code",
    prac_name = "Practice name",
    ccg_id = "Sub-ICB Location code",
    name= "Sub-ICB Location name",
    regional_team_id = "Region code",
    region = "region name") %>%
  tab_options(table.font.size = 10,
  data_row.padding = px(1)) %>% 
    tab_style(
      style = cell_borders(
      sides = "all",
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = list(cells_body(), cells_column_labels())) 

```

```{r}
#| label: 20mg-filter-all-months
#| include: false

prac_count_20<-df_atorvastatin_prac %>% 
    filter(str_detect(bnf_name, '20')) %>%
    group_by(practice,month) %>% 
    summarise(practice =last(practice)) %>%
    count(practice) %>%
    filter(n == study_months)

dropped_20 <- prac_count_all_months %>%
  anti_join(prac_count_20)

```

`{r} label_number_n(nrow(prac_count_all_months) - nrow(prac_count_20))` of the `{r} label_number_n(nrow(prac_count_all_months))` practices practices which contribute to every month are lost when restricting to practices that prescribe 20mg atorvastatin.

```{r}
#| label: table-20mg-all-months-filter
#| echo: false
#| warning: false

## table of practices lost when filtered to 20mg
df_dropped_20 <- dropped_20 %>%
  left_join(df_stats_prac_full) %>%
  group_by(practice) %>%
  summarise(prac_name = last(name),
            ccg_id=last(ccg_id)) %>%
  left_join(df_hscic_ccgs,by=join_by(ccg_id ==code)) %>%
  left_join(df_regions,by=join_by(regional_team_id ==regional_team )) %>%
  select(practice,prac_name,ccg_id,name,regional_team_id,region) %>%
  ungroup()

gt(df_dropped_20) %>%
  sub_missing(missing_text = "-") %>%
    cols_label(
    practice = "Practice code",
    prac_name = "Practice name",
    ccg_id = "Sub-ICB Location code",
    name= "Sub-ICB Location name",
    regional_team_id = "Region code",
    region = "region name") %>%
  tab_options(table.font.size = 10,
  data_row.padding = px(1)) %>% 
    tab_style(
      style = cell_borders(
      sides = "all",
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = list(cells_body(), cells_column_labels())) 

```

#### Sub-ICB Locations

```{r}
#| label: sicbl_compare
#| echo: false
#| warning: false
#### pratice per month (combines all items per practice per month)
normalised_prescribing_pct <- df_atorvastatin_prac %>%
  select(month,practice,pct)

practice_statistics_pct_id <- df_hscic_practice_statistics %>%
  select(month,practice,pct_id)

practices_ccg_id <- df_hscic_practices %>%
  select(code,ccg_id)

sicbl_comp <- normalised_prescribing_pct %>%
  left_join(practice_statistics_pct_id) %>%
  left_join(practices_ccg_id, by = join_by(practice==code)) %>%
  mutate(pct_pctid=pct==pct_id,
         pct_ccgid=pct==ccg_id,
         pctid_ccgid=pct_id==ccg_id,
  )


id_sicbl_diff<-sicbl_comp %>%
  filter(pctid_ccgid==FALSE) %>%
  group_by(ccg_id) %>%
  arrange(month) %>%
  summarise(ccg_id=last(ccg_id)) 

id_sicbl_diff_prac<-sicbl_comp %>%
  filter(pctid_ccgid==FALSE) %>%
  group_by(practice) %>%
  arrange(month) %>%
  summarise(practice=last(practice)) 

sicbl_diff<-sicbl_comp %>%
  right_join(id_sicbl_diff) %>% 
  group_by(ccg_id,pct_id) %>%
  arrange(month) %>%
  summarise(ccg_id=last(ccg_id),
            pct_id=last(pct_id))

tab_sicbl_diff_practice <- sicbl_comp %>%
  right_join(id_sicbl_diff_prac) %>% 
  group_by(practice,pct_id) %>%
  arrange(month) %>%
  summarise(practice=last(practice),
            pct_id=last(pct_id)) %>%
  left_join(df_hscic_practices %>% select(code,name),by=join_by(practice==code)) %>%
  left_join(df_hscic_ccgs %>% select(code,name),by=join_by(pct_id==code)) %>%
  replace_na(list(pct_id="Missing code",name.y="Missing code")) %>%
  mutate(counter = row_number(practice),
  direction = case_when(counter ==1 ~ "Original",
  counter ==2 ~ "1st_change",
  counter==3 ~ "2nd_change")) %>%
  ungroup() %>%
  reframe(practice_code=practice,
            practice_name=name.x,
            SICBL_code=pct_id,
            SICBL_name=name.y,
            direction) %>%
  pivot_wider(names_from = direction, values_from = c(SICBL_code,SICBL_name),values_fill = " ") %>%
  ungroup()

```

pct and ccg_id always agree. `{r} nrow(id_sicbl_diff_prac)` from `{r} nrow(id_sicbl_diff)` sub-icb locations have changed sub-icb location Appendix 1.

#### Practices which changed region during the study period

```{r}
#| label: compare-to-monthly
#| echo: false
#| warning: false
#### pratice per month (combines all items per practice per month)
df_atorvastatin_prac_month <- df_atorvastatin_prac %>%
  left_join(df_stats_prac) %>%
  group_by(month,practice) %>%
  summarise(
            items=sum(items),
            total_list_size=mean(total_list_size, na.rm = TRUE),
            regional_team = last(regional_team ),
            ccg_id = last(ccg_id),
            rate = items / total_list_size *1000)  %>%
  ungroup() %>%
  mutate(month = as.Date(month))


### compare practice to region dfs
### 
df_atorvastatin_chem <- df_atorvastatin %>%
  # filter(str_detect(bnf_name, '20')) %>%
  group_by(month,regional_team) %>%
  summarise(items=sum(items),
            total_list_size=mean(total_list_size),
            rate = items / total_list_size *1000)  %>%
  ungroup() 



df_atorvastatin_prac_compare <- df_atorvastatin_prac_month %>%
  ungroup() %>%
  group_by(month,regional_team) %>%
  summarise(items=sum(items),
            total_list_size=sum(total_list_size, na.rm = TRUE),
            rate = items / total_list_size *1000) %>%
  select(month,regional_team,items,total_list_size,rate) %>%
  ungroup() %>%
  full_join(df_atorvastatin_chem,by=join_by(month,regional_team)) %>%
  mutate(items= items.x-items.y,
         total_list_size = total_list_size.x-total_list_size.y)
       ##### difference in list size not in items ### list sometimes more / sometimes less

if( min(df_atorvastatin_prac_compare$items,na.rm = T) != 0 | max(df_atorvastatin_prac_compare$items,na.rm = T) ){ print("Items are not equal between regional and practice data")}

```

Sub-ICB locations (SICBL) can be identified via 3 different columns in three different datasets. The ccg_id column in hscic.practices and the pct column in hscic.normalised-prescribing are static and show the SICBL as it is now. The pct_id column in hscic.practice-statistics shows the SICBL as it was at the time.

The difference in population between the `pct_id` and `ccg_id` definition of SICBL can be seen when plotting the list size per month for each region. In the `pct_id` plot there is a dip in list size in the North West in June 2022, corresponding to a rise in Midlands.

```{r}
#| label: figure-list-size-season-pct
#| echo: false
#| warning: false
#| fig.width: 30
#| fig.height: 10


### season plot list sizes
# 

regions<-unique(df_regions$regional_team)
  
season_pct<-df_stats_prac %>%
  left_join(df_hscic_ccgs,by=join_by(pct_id ==code)) %>% 
  group_by(month,regional_team_id) %>%
  summarise(
    total_list_size=sum(total_list_size, na.rm = TRUE))  %>%
  left_join(df_regions,by=join_by(regional_team_id ==regional_team )) %>%
  ungroup() %>%
  arrange(month)

for(i in regions){
  
  
  assign(glue("season_plot_total_list_size_{df_regions %>% filter(regional_team == i) %>% select(region)}"),   assign(glue("season_plot_{i}"), ggseasonplot(ts(season_pct %>% filter(regional_team_id == i) %>% select(total_list_size), start = c(2020,01), frequency = 12),main = glue("{df_regions %>% filter(regional_team == i) %>% select(region)} list size seasonal plot"))))
  
}

seasons_combined_list_size_pct <- patchwork::wrap_plots(`season_plot_total_list_size_East of England`,`season_plot_total_list_size_London`,`season_plot_total_list_size_Midlands`, `season_plot_total_list_size_North East and Yorkshire`,
                                                    `season_plot_total_list_size_North West`,`season_plot_total_list_size_South East`,`season_plot_total_list_size_South West`)


ggsave(
  filename = here::here(
    "output",
    "data_check",
    "atorvastatin",
    glue("list_size_seasons_pct.png")),
  seasons_combined_list_size_pct,
  dpi = 600,
  width = 40,
  height = 15,
  units = "cm"
)

seasons_combined_list_size_pct

```

However, we do not see the same dip in the `ccg_id` plot.

```{r}
#| label: figure-list-size-season-sicbl
#| echo: false
#| warning: false
#| fig.width: 30
#| fig.height: 10



### season plot list sizes
# 
 
season_sicbl<-df_stats_prac %>%
  left_join(df_hscic_ccgs,by=join_by(ccg_id ==code)) %>% 
  group_by(month,regional_team_id) %>%
  summarise(
            total_list_size=sum(total_list_size, na.rm = TRUE))  %>%
  left_join(df_regions,by=join_by(regional_team_id ==regional_team )) %>%
  ungroup() %>%
  arrange(month)



for(i in regions){
  
  
  assign(glue("season_plot_total_list_size_{df_regions %>% filter(regional_team == i) %>% select(region)}"),   assign(glue("season_plot_{i}"), ggseasonplot(ts(season_sicbl %>% filter(regional_team_id == i) %>% select(total_list_size), start = c(2020,01), frequency = 12),main = glue("{df_regions %>% filter(regional_team == i) %>% select(region)} list size seasonal plot"))))
  
}

seasons_combined_list_size_sicbl <- patchwork::wrap_plots(`season_plot_total_list_size_East of England`,`season_plot_total_list_size_London`,`season_plot_total_list_size_Midlands`, `season_plot_total_list_size_North East and Yorkshire`,
                                                    `season_plot_total_list_size_North West`,`season_plot_total_list_size_South East`,`season_plot_total_list_size_South West`)



ggsave(
  filename = here::here(
    "output",
    "data_check",
    "atorvastatin",
    glue("list_size_seasons_sibl.png")),
  seasons_combined_list_size_sicbl,
  dpi = 600,
  width = 40,
  height = 15,
  units = "cm"
)

seasons_combined_list_size_sicbl


```

```{r}
#| label: ccg-region-change
#| echo: false
#| warning: false

### practice region from CCG (sicbl)
df_stats_prac_region_ccg <- df_stats_prac %>%
  left_join(df_hscic_ccgs,by=join_by(ccg_id ==code)) 

### sicbl which change region (pct_id)
sicbl_change<-df_stats_prac_region_ccg %>% 
  group_by(pct_id,regional_team_id) %>%
  summarise(pct_id = last(pct_id),
            regional_team_id = last(regional_team_id)) %>% 
  count(pct_id) %>% 
  filter(n>1)
                
```

```{r}
#| label: table-ccg-region-change
#| echo: false
#| warning: false

##### ignore this. Multiple practices in sicbl some change some don't
# sicbl_change_regions <- sicbl_change %>%
#   left_join(df_stats_prac_region_ccg %>% 
#               group_by(pct_id,regional_team_id) %>%
#               arrange(month) %>%
#               summarise(pct_id = last(pct_id),
#                         regional_team_id = last(regional_team_id),#
#                         month = last(month))) %>%
#   arrange(pct_id, month) %>%
#   mutate(counter = row_number(pct_id),
#          direction = case_when(counter ==1 ~ "From",
#                                counter ==2 ~ "To")) %>%
#   left_join(df_regions,by = join_by(regional_team_id ==regional_team)) %>%
#   select(pct_id,direction,region) %>%
#   pivot_wider(names_from = direction, values_from = region) %>%
#   ungroup()

# gt(sicbl_change_regions)
```

```{r}
#| label: practice-region-change
#| echo: false
#| warning: false

### practice region from CCG (sicbl)
df_stats_prac_region_pct_id <- df_stats_prac %>%
  left_join(df_hscic_ccgs,by=join_by(pct_id ==code)) 

### practices which change region (pct_id)
practice_change<-df_stats_prac_region_pct_id %>% 
  group_by(practice,regional_team_id) %>%
  summarise(practice = last(practice),
            regional_team_id = last(regional_team_id)) %>% 
  count(practice) %>% 
  filter(n>1)

practice_change_regions <- practice_change %>%
    left_join(df_stats_prac_region_pct_id %>% 
  group_by(practice,regional_team_id) %>%
  arrange(month) %>%
  summarise(practice = last(practice),
            regional_team_id = last(regional_team_id),
            stp_id = last(stp_id),
            sicbl_name = last(name.y),
            pct_id =last(pct_id),
            month = last(month))) %>%
                arrange(practice, month) 

### df_hscic_stp previously loaded from prac_stps.csv 
tab_practice_change_regions <-  practice_change_regions  %>%
                mutate(counter = row_number(practice),
                direction = case_when(counter ==1 ~ "From",
                counter ==2 ~ "To")) %>%
                left_join(df_regions,by = join_by(regional_team_id ==regional_team)) %>%
                left_join(df_hscic_stp,by=join_by(stp_id==code)) %>%
                select(practice,direction,region,name,sicbl_name,pct_id,month) %>%
                pivot_wider(names_from = direction, values_from = c(region,name,sicbl_name,pct_id,month)) %>%
                select(-month_To) %>%
  ungroup()              
                
```

There were `{r} length(tab_practice_change_regions$practice)` practices, from `{r} length(unique(tab_practice_change_regions$sicbl_name_To))` Integrated Care Boards (ICB) and `{r}  nrow(sicbl_change)` sub-ICB locations, which changed region during the study period.

```{r}
#| label: table-practice-region-change
#| echo: false
#| warning: false
               


gt(tab_practice_change_regions) %>%
  cols_label(
    name_From = "from",
    sicbl_name_From = "from",
    pct_id_From = "from",
    region_From = "from",
    name_To = "to",
    sicbl_name_To = "to",
    pct_id_To = "to",
    region_To = "to",
    month_From = "Date of change"
  ) %>%
     tab_spanner(
    label = "ICB",
    columns = starts_with("name")
  )  %>%
     tab_spanner(
    label = "SICBL",
    columns = starts_with("sicbl_name")
  )  %>%
     tab_spanner(
    label = "SICBL code",
    columns = starts_with("pct_id")
  )  %>%
     tab_spanner(
    label = "Region",
    columns = starts_with("region")
  ) %>%
  tab_options(table.font.size = 10,
  data_row.padding = px(1)) %>% 
    tab_style(
      style = cell_borders(
      sides = "all",
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = list(cells_body(), cells_column_labels())) 
```

In order to investigate if the practices change list size when they move region I plotted the list size by month.

```{r}
#| label: figure-practice-region-change
#| echo: false
#| warning: false
               
id_practice_change_regions <-  practice_change_regions  %>%
                mutate(counter = row_number(practice)) %>%
                filter(counter=="1") %>%
                select(practice,month) %>% 
                mutate(month =as.Date(month)) %>% 
  ungroup()


for(prac in id_practice_change_regions$practice){
     assign(glue("plot_practice_change_regions_{prac}"),df_atorvastatin_prac_region %>% 
           filter(practice==prac) %>%
           ggplot(aes(x = month, y = total_list_size)) +
           geom_line() +
           scale_x_date(date_breaks = "4 months",limits = c(start_date_ator,end_date_ator))+
           scale_y_continuous( limits = c(0, NA),labels = label_comma()) +
           theme(axis.text.x = element_text(angle =90)) +
           geom_vline(xintercept= as.Date(id_practice_change_regions %>%  
                        filter(practice==prac) %>% 
                        select(month) %>%
                        unlist())) +
                        ggtitle(glue("{prac} {df_stats_prac %>% group_by(practice) %>% summarise(name=last(name)) %>% filter(practice ==prac)%>%select(name) %>% unlist()}")))
}

id_practice_change_regions <- id_practice_change_regions %>%
  mutate(get=glue("plot_practice_change_regions_{practice}"))

plot_practice_change_regions <- patchwork::wrap_plots(lapply(id_practice_change_regions$get,get),ncol =3)

  ggsave(
    filename = here::here(
      "output",
      "data_check",
      "atorvastatin",
      "practice_change_region",
      glue("practice_change_region_{prac}.png")),
    plot_practice_change_regions,
    dpi = 600,
    width = 40,
    height = 30,
    units = "cm",
    create.dir = T
  )



plot_practice_change_regions


           
```

Although the practices all had changes in list size after the change in region this could be seen as a continuation of the pattern seen before the practice change, rather than evidence of the practice merging with another.

#### Practices that initiate prescribing

```{r}
#| label: miss-start
#| echo: false
#| warning: false

miss_start<-unique(df_atorvastatin_prac_region %>% 
    filter(month=="2020-01-01") %>% 
    select(practice))


id_miss_start<-df_atorvastatin_prac_region %>%
    anti_join(miss_start) %>%
    group_by(practice) %>%    
    count(NAs = is.na(total_list_size)) %>%  ### remove practices that are always missing listsize
    arrange(practice,NAs) %>%
    filter(NAs==F) %>%
    select(practice) 

  miss_start_sas <- setting_sas %>%
    inner_join(id_miss_start) %>%
    select(practice) %>%
    unlist()

other_atypical <- df_setting %>%
  anti_join(setting_sas)

miss_start_other_atypical <- other_atypical %>%
 inner_join(id_miss_start) %>%
 select(practice) %>%
 unlist()

miss_start_just_normal_practices <- id_miss_start %>%
 anti_join(df_setting) %>%
 select(practice) %>%
 unlist()

atypical_initiators <- df_setting %>% 
  inner_join(id_miss_start)

```

I defined practices that initiate prescribing as practices with no atorvastatin prescribing at the start of the study period. There were `{r} nrow(id_miss_start)` practices that initiated prescribing during the study period. In order to investigate if there is any pattern to opening practices I plotted their list size per month.

`{r} length(miss_start_sas)` of these practices are SAS practices and `{r} length(miss_start_other_atypical)` of these practices are specialist, urgent, asylum, pilot or nursing home practices.

```{r}
#| label: table-miss_start_atypical
#| echo: false
#| warning: false
atypical_initiators %>% 
  gt() %>%
  cols_label(highest_list_size = "maximum list size") %>%
  tab_options(table.font.size = 10,
  data_row.padding = px(1)) %>%
    tab_style(
      style = cell_borders(
      sides = "all",
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = list(cells_body(), cells_column_labels())) %>% tab_style(
      style = cell_borders(
      sides = "all",
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = list(cells_body(), cells_column_labels())) 
```

The remaining `{r} length(miss_start_just_normal_practices)` practices appear to be correctly categorised as GP practices.

```{r}
#| label: miss_start_just_normal_practices
#| echo: false
#| warning: false
for(prac in miss_start_just_normal_practices){
  assign(glue("plot_opening_practices_{prac}"), df_atorvastatin_prac_region %>% 
           filter(practice==prac) %>%
           ggplot(aes(x = month, y = total_list_size)) +
           geom_line() +
           scale_x_date(date_breaks = "4 months",limits = c(start_date_ator,end_date_ator))+
           scale_y_continuous( limits = c(0, NA),labels = label_comma()) +
           theme(axis.text.x = element_text(angle =90)) +
                        ggtitle(glue("{prac} {df_stats_prac %>% group_by(practice) %>% summarise(name=last(name)) %>% filter(practice ==prac)%>%select(name) %>% unlist()}")) )
}


just_normal_practices <- id_miss_start %>%
 anti_join(df_setting) %>%
 select(practice) %>%
 mutate(get=glue("plot_opening_practices_{practice}"))

plot_opening_practices <- patchwork::wrap_plots(lapply(just_normal_practices$get,get),ncol =3)


  ggsave(
    filename = here::here(
      "output",
      "data_check",
      "atorvastatin",
      "opening_practices",
      glue("opening_practices_combined.png")),
    plot_opening_practices,
    dpi = 600,
    width = 19,
    height = 25,
    units = "cm",
    create.dir = T
  )


```

There was no single pattern seen for practices that initiated prescribing during the study period. Several practices opened at, or close to, their highest list size. We also see practices that started with a low list size and grew their number of patients over time.

#### Practices that stop prescribing

```{r}
#| label: miss-end
#| echo: false
#| warning: false

miss_end <- unique(df_atorvastatin_prac_region %>% 
    filter(month=="2025-07-01") %>% 
    select(practice))

id_miss_end<-df_atorvastatin_prac_region %>%
    anti_join(miss_end) %>%
    group_by(practice) %>%    
    count(NAs = is.na(total_list_size)) %>%  ### remove practices that are always missing listsize
    arrange(practice,NAs) %>%
    filter(NAs==F) %>%
    select(practice) 

 atypical_end <- id_miss_end %>% 
  inner_join(df_setting) %>%
  ungroup()

id_atypical_end <- atypical_end %>% 
  select(practice) %>% 
  unlist()

id_miss_end <- id_miss_end %>%
  unlist()
```

Similarly, I defined practices that stop prescribing as practices with no atorvastatin prescribing at the end of the study period. There were `{r} length(id_miss_end)` practices that stopped prescribing during the study period. In order to investigate if there is any pattern to these practices I plotted their list size per month.

`{r} length(id_atypical_end)` of the closing practices were atypical.

```{r}
#| label: table-miss_end_atypical
#| echo: false
#| warning: false
atypical_end %>% 
  gt() %>%
  cols_label(highest_list_size = "maximum list size") %>%
  tab_options(table.font.size = 10,
  data_row.padding = px(1)) %>%
    tab_style(
      style = cell_borders(
      sides = "all",
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = list(cells_body(), cells_column_labels())) %>% tab_style(
      style = cell_borders(
      sides = "all",
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = list(cells_body(), cells_column_labels())) 
```

```{r}
#| label: plot-miss-end
#| echo: false
#| warning: false
#| fig.width: 30
#| fig.height: 10

for(prac in id_miss_end){
     assign(glue("plot_closing_practices_{prac}"),df_atorvastatin_prac_region %>% 
           filter(practice==prac) %>%
           ggplot(aes(x = month, y = total_list_size)) +
           geom_line() +
           scale_x_date(date_breaks = "4 months",limits = c(start_date_ator,end_date_ator)) +
           scale_y_continuous( limits = c(0, NA),labels = label_comma()) +
           theme(axis.text.x = element_text(angle =90)) +
                        ggtitle(glue("{prac} {df_stats_prac %>% group_by(practice) %>% summarise(name=last(name)) %>% filter(practice ==prac)%>%select(name) %>% unlist()}"))) 

  ggsave(
    filename = here::here(
      "output",
      "data_check",
      "atorvastatin",
      "closing_practices",
      glue("closing_practices_{prac}.png")),
    get(glue("plot_closing_practices_{prac}")),
    dpi = 600,
    width = 40,
    height = 15,
    units = "cm",
    create.dir = T
  )

}

```

The remaining `{r} length(id_miss_end) - length(id_atypical_end)` practices can be into 4 different patterns:

A sharp decrease in list size before closing

```{r}
#| label: plot-sharp-decrease
#| echo: false
#| warning: false
#| fig.width: 30
#| fig.height: 10

plot_closing_practices_E82072

```

Slow decrease in list size before closing

```{r}
#| label: plot-slow-decrease
#| echo: false
#| warning: false
#| fig.width: 30
#| fig.height: 10

plot_closing_practices_Y00110

```

A period of low list size before closing

```{r}
#| label: plot-long-tail
#| echo: false
#| warning: false
#| fig.width: 30
#| fig.height: 10

plot_closing_practices_C83653
```

No decrease in list size before closing

```{r}
#| label: plot-no-decrease
#| echo: false
#| warning: false
#| fig.width: 30
#| fig.height: 10

plot_closing_practices_A81043
```

Activity after closing

```{r}
#| label: plot-continued-activity
#| echo: false
#| warning: false
#| fig.width: 30
#| fig.height: 10

plot_closing_practices_F86642


 file.copy(file.path(here::here(
     "output",
      "data_check",
    "atorvastatin",
    "closing_practices"),
    c("closing_practices_E82072.png","closing_practices_Y00110.png","closing_practices_C83653.png","closing_practices_A81043.png","closing_practices_F86642.png")), here::here(  "output",   "data_check",  "atorvastatin",    "closing_practices","examples"))


```

### Discussion

#### Summary

#### Limitations

This study has several limitations. First, the study investigates practices which prescribe atorvastatin. Although atorvastatin is a widely prescribed drug it may not be representative of wider prescribing behaviour. Second, the study tried to identify atypical practices based on name alone. Although this strategy is demonstrative of the issues of atypical practice a more robust strategy would be required to fully identify atypical practices. Third, this study did not investigate practices which merge during the study period. Fourth, although this study highlighted examples of practices with missing list size and discussed some of the potential issues contributing to missing list size, it did not investigate every practice with missing list size information.

#### Conclusion

This report aims to aid the use of OpenPrescribing data. OpenPrescribing collates datasets from the NHS Business Services Authority. There are several issues that should be carefully considered before using datasets from the NHS Business Services Authority. The decisions on how each of these issues should be addressed is highly dependent on the individual study question. Although this report focuses primarily on atorvastatin prescribing the findings can be used to inform decision making on other studies using either OpenPrescribing or NHS Business Services Authority data.

Appendix 1

```{r}
#| label: table-sicbl_diff_practice
#| echo: false
#| warning: false
               


gt(tab_sicbl_diff_practice) %>%
  cols_label(
    practice_code = "Practice code",
    practice_name = "Practice name",
    SICBL_code_Original = "Original SICBL code",
    SICBL_code_1st_change = "2nd SICBL code",
    SICBL_code_2nd_change = "3rd SICBL code",
    SICBL_name_Original = "Original SICBL name",
    SICBL_name_1st_change = "2nd SICBL name",
    SICBL_name_2nd_change = "3rd SICBL name"
  )  %>%
  tab_options(table.font.size = 10,
  data_row.padding = px(1)) %>% 
    tab_style(
      style = cell_borders(
      sides = "all",
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = list(cells_body(), cells_column_labels())) 
```

Appendix 2 `{r} list_sizeless`